---
title: "Untitled"
author: "Emily Zhou"
date: "2023-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


- **The EMS dataset** that includes all responses to heroin-related overdose incidents from the Cincinnati Fire Department from 2015-2023. 
https://data.cincinnati-oh.gov/Safety/Cincinnati-Fire-Incidents-CAD-including-EMS-ALS-BL/vnsz-a3wp

- **ACS data**: adult population size, the population groups by age, the ratio of gender, and the ratio of race/ethnicity, education, income, and poverty.

- **Hospital data** from Homeland Infrastructure Foundation-Level Data (HIFLD)
 https://hifld-geoplatform.opendata.arcgis.com/maps/geoplatform::hospitals/about.

- **Substance Abuse and Mental Health Services Administration (SAMHSA) data** geocoded using google
https://dpt2.samhsa.gov/treatment/directory.aspx

- **Crime rate data** from the Cincinnati Police Department from 2016 to 2022
https://data.cincinnati-oh.gov/safety/PDI-Police-Data-Initiative-Crime-Incidents/k59e-2pvf

- **311 complaints** non-emergency service requests from 2016 to 2021
https://data.cincinnati-oh.gov/Thriving-Neighborhoods/Cincinnati-311-Non-Emergency-Service-Requests/4cjh-bm8b

-**built environment data** all from OSM

- **Cincinnati boundary**
https://data-cagisportal.opendata.arcgis.com/datasets/ed78f4754b044ac5815d0a9efe9bb336_1/explore

```{r heroin data}
library(tidyverse)
library(sf)
library(lubridate)
library(here)

heroin <- read.csv(here("data", "Cincinnati_Fire_Incidents__CAD___including_EMS__ALS_BLS_.csv"))
cincinnati <- st_read(here("data", "Cincinnati_City_Boundary.geojson")) %>% st_transform("EPSG:3735")

heroin <- heroin %>% 
  filter(CFD_INCIDENT_TYPE_GROUP != "NON-PROTOCOL PROBLEM TYPES") %>% 
  filter(is.na(LATITUDE_X) == FALSE & is.na(LONGITUDE_X) == FALSE) %>%  # remove incidents withouth spatial info
  filter(is.na(DISPOSITION_TEXT) == FALSE         # remove record with un-associated disposition codes
         & CFD_INCIDENT_TYPE_GROUP !="CN: CANCEL" # remove records from canceled calls, false alarm, and duplicated calls
         & CFD_INCIDENT_TYPE_GROUP !="CANCEL INCIDENT" 
         & CFD_INCIDENT_TYPE_GROUP !="CN: CANCEL,DEF: DEFAULT"
         & CFD_INCIDENT_TYPE_GROUP !="CN: CANCEL,EMSF: FALSE"
         & CFD_INCIDENT_TYPE_GROUP !="CN: CANCEL,DUPF: DUPLICATE"
         & CFD_INCIDENT_TYPE_GROUP !="CN: CANCEL,FALA: FIRE FALSE AC"
         & CFD_INCIDENT_TYPE_GROUP !="CN: CANCEL,MEDD: MT DISREGARDE"
         & CFD_INCIDENT_TYPE_GROUP !="DUPF: DUPLICATE"
         & CFD_INCIDENT_TYPE_GROUP !="DUPLICATE INCIDENT"
         & CFD_INCIDENT_TYPE_GROUP !="MAL: SYSTEM MALFUNCTION") %>% 
  mutate(time = mdy_hms(CREATE_TIME_INCIDENT, tz = "UTC"),  
         year_column = year(time)) %>% 
  st_as_sf(., coords = c("LONGITUDE_X", "LATITUDE_X"), crs = 4326) %>% 
  st_transform("EPSG:3735") %>% 
  st_intersection(cincinnati %>% select(OBJECTID),.)  # remove incidents outside of study area


```


```{r}

ggplot()+
  geom_sf(data=cincinnati) +
  geom_sf(data = heroin)

```


```{r hospital data}

hospitals <- st_read(here("data", "Hospitals.geojson"))

hospitals <- hospitals %>% 
  filter(STATE == "OH" & CITY == "CINCINNATI") %>% 
  st_transform("EPSG:3735")# select those only in Cincinnati

```


Specific crimes were selected for inclusion in analyses based on having been identified in the literature as a spatial risk factor for drug overdose or substance use behavior or for having a theoretical relationship with drug overdose. 

```{r crime data}

crime <- read.csv(here("data", "PDI__Police_Data_Initiative__Crime_Incidents.csv"))
  
crime_list <- c("AGGRAVATED ASSAULT", "BURGLARY", "BREAKING AND ENTERING", "ROBBERY", "DOMESTIC VIOLENCE", "MURDER	", "RAPE", "THEFT")

crime <- crime %>% 
  filter(OFFENSE %in% crime_list) %>% 
  filter(is.na(LATITUDE_X) == FALSE & is.na(LONGITUDE_X) == FALSE) %>%
  st_as_sf(., coords = c("LONGITUDE_X", "LATITUDE_X"), crs = 4326) %>% 
  st_transform("EPSG:3735") %>% 
  mutate(time = mdy_hms(DATE_REPORTED, tz = "UTC"),  
         year = year(time))

```

`
```{r 311 complaints}

complaints <- read_csv(here("data", "Cincinnati_311__Non-Emergency__Service_Requests.csv"))
  
complaints <- complaints %>% 
  filter(is.na(LATITUDE) == FALSE & is.na(LONGITUDE) == FALSE) %>%
  st_as_sf(., coords = c("LONGITUDE", "LATITUDE"), crs = 4326) %>% 
  st_transform("EPSG:3735") %>% 
  mutate(time = mdy_hms(REQUESTED_DATE, tz = "UTC"),  
         year = year(time))

```

The connection between substance use and built environment variables (access to public restrooms, access to pharmacies, and driving distance to services, defined in our study as fast-food restaurants, gas stations, and highway exits) are also important. Public restrooms are associated with people who inject drugs (PWID) because many people (one study estimates 48%) use drugs in these spaces.Pharmacies represent an important access variable for several reasons. During the initial wave of the overdose crisis, pharmaceutical prescriptions, either legitimate, diverted, or potentially inappropriate, fed the opioid supply. In addition, naloxone (a medication to reverse an opioid overdose) is available at pharmacies without a prescription, although this provision may vary by neighborhood socio-demographic levels.

```{r built environment data}

# gas station
fuel <- st_read(here("data", "fuel.geojson")) %>% st_transform("EPSG:3735")

# fast food restaurant 
fastfood <- st_read(here("data", "fastfood.geojson")) %>% st_centroid() %>% st_transform("EPSG:3735")

# public parks
parks <- st_read(here("data", "parks.geojson")) %>% st_transform("EPSG:3735") %>% st_centroid()

# pharmacies
pharmacy <- st_read(here("data", "pharmacy.geojson")) %>% st_transform("EPSG:3735") %>% st_centroid()

```


```{r census data}

library(tidycensus)
tidycensus::census_api_key("e79f3706b6d61249968c6ce88794f6f556e5bf3d", overwrite = TRUE)

cincinnati20 <- get_acs(geography = "tract", 
          variables = c(
            "B01001_001E", # total population
            "B01001_002E", # total male
            "B01001_011E", # male 25-29
            "B01001_012E",
            "B01001_013E",
            "B01001_014E",
            "B01001_015E",
            "B01001_016E", # male 50-54
            "B01001_026E", # total female
            "B01001_035E", # female 25-29
            "B01001_036E",
            "B01001_037E",
            "B01001_038E",
            "B01001_039E",
            "B01001_040E", # female 50-54
            "B02001_002E", # white population
            "B02001_003E", # black population
            "B02001_005E", # asian population
            "B03002_012E", # latinx population
            "B19013_001E", # median household income
            "B06012_002E", # poverty
            "B06009_005E" # bachelor
            ), 
          year=2020, state="OH", county="Hamilton", 
          geometry=TRUE, output="wide") %>%
  st_transform("EPSG:3735")

cincinnati_tracts <- st_read(here("data", "cincinnati_tracts.geojson")) %>% st_transform("EPSG:3735")

cincinnati_tracts <- cincinnati_tracts %>% st_drop_geometry() %>% dplyr::select(GEOID) %>% as.list(GEOID)

cincinnati20 <- cincinnati20 %>% 
  filter(GEOID %in% cincinnati_tracts$GEOID) %>% 
  rename(Totalpop = B01001_001E,
         MedHHInc = B19013_001E) %>% 
  mutate(pop25_54 = B01001_011E + B01001_012E + B01001_013E + B01001_014E + B01001_015E + B01001_016E + B01001_035E + B01001_036E + B01001_037E + B01001_038E+ B01001_039E + B01001_040E) %>% 
  mutate(MF_ratio = B01001_002E / B01001_026E) %>% 
  mutate(race_ratio = B02001_002E / (B02001_003E + B02001_005E + B03002_012E)) %>% 
  mutate(pctPoverty = B06012_002E / Totalpop) %>% 
  mutate(pctBachelor = B06009_005E / Totalpop) %>% 
  dplyr::select(pop25_54, Totalpop, MedHHInc, GEOID, MF_ratio, race_ratio, pctPoverty, pctBachelor)


```


```{r}

rehab <- read.csv(here("data", "rehabilitation.csv"))
  
rehab <- rehab %>% 
  filter(is.na(Latitude) == FALSE & is.na(Longitude) == FALSE) %>%
  st_as_sf(., coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform("EPSG:3735")
  

```

heroin

Predictor Variables so far: 
- hospitals
- crime rate (2016-2018)
- 311 complaints (2016-2018)
- total populations
- population between 25-54
- gender ratio
- white over minority ratio
- education
- median household income
- poverty
- gas station
- fast food restaurant
- public parks
- pharmacies
- rehabilitation centers

```{r select certain data}

heroin <- heroin %>% 
  filter(year_column %in% c(2016, 2017, 2018, 2019, 2020))

crime <- crime %>% 
  filter(year %in% c(2016, 2017, 2018, 2019, 2020))

complaints <- complaints %>% 
  filter(year %in% c(2016, 2017, 2018, 2019, 2020))

# split into training and testing by year here?

```


```{r}

fishnet <- st_make_grid(cincinnati,
               cellsize = 820, 
               square = TRUE) %>%
  .[cincinnati] %>%           
  st_sf() %>%
  mutate(uniqueID = 1:n())

ggplot() +
  geom_sf(data=fishnet, color="black", fill="#FFF5EE") +
  labs(title = "Fishnet of Cincinnati") +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks =element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.subtitle = element_text(size = 9,face = "italic"),
        plot.title = element_text(size = 12, face = "bold"),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth=0.8)
        )
```


```{r}

heroinTrain <- heroin %>% 
  filter(year_column %in% c(2016, 2017, 2018))
heroin19 <- heroin %>% 
  filter(year_column == 2019)
heroin20 <- heroin %>% 
  filter(year_column == 2020)

crimeTrain <- crime %>% 
  filter(year %in% c(2016, 2017, 2018))
crime19 <- crime %>% 
  filter(year == 2019)
crime20 <- crime %>% 
  filter(year == 2020)

complaintTrain <- complaints %>% 
  filter(year %in% c(2016, 2017, 2018))
complaint19 <- complaints %>% 
  filter(year == 2019)
complaint20 <- complaints%>% 
  filter(year == 2020)

# add heroin in net
heroinTrain_net <- 
  dplyr::select(heroinTrain) %>%  
  mutate(countHeroin = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countHeroin = replace_na(countHeroin, 0),
         uniqueID = 1:n(), 
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

# adding crime to net
heroinTrain_net <- heroinTrain_net %>% 
  st_join(crimeTrain, ., join=st_within) %>% 
  st_drop_geometry() %>%
  group_by(uniqueID) %>%
  summarize(Crimecount = n()) %>%
    left_join(heroinTrain_net, . ) %>%
    st_sf() %>%
  mutate(Crimecount = ifelse(is.na(Crimecount), 0, Crimecount))

# add 311 to net
heroinTrain_net <- heroinTrain_net %>% 
  st_join(complaintTrain, ., join=st_within) %>% 
  st_drop_geometry() %>%
  group_by(uniqueID) %>%
  summarize(Complaintscount = n()) %>%
    left_join(heroinTrain_net, . ) %>%
    st_sf() %>%
  mutate(Complaintscount = ifelse(is.na(Complaintscount), 0, Complaintscount))

```


```{r}

# add 19 materials
heroin19_net <- 
  dplyr::select(heroin19) %>%  
  mutate(countHeroin = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countHeroin = replace_na(countHeroin, 0),
         uniqueID = 1:n(), 
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

heroin19_net <- heroin19_net %>% 
  st_join(crime19, ., join=st_within) %>% 
  st_drop_geometry() %>%
  group_by(uniqueID) %>%
  summarize(Crimecount = n()) %>%
    left_join(heroin19_net, . ) %>%
    st_sf() %>%
  mutate(Crimecount = ifelse(is.na(Crimecount), 0, Crimecount))


heroin19_net <- heroin19_net %>% 
  st_join(complaint19, ., join=st_within) %>% 
  st_drop_geometry() %>%
  group_by(uniqueID) %>%
  summarize(Complaintscount = n()) %>%
    left_join(heroin19_net, . ) %>%
    st_sf() %>%
  mutate(Complaintscount = ifelse(is.na(Complaintscount), 0, Complaintscount))
```

```{r}

# add 20 material
heroin20_net <- 
  dplyr::select(heroin20) %>%  
  mutate(countHeroin = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countHeroin = replace_na(countHeroin, 0),
         uniqueID = 1:n(), 
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

heroin20_net <- heroin20_net %>% 
  st_join(crime20, ., join=st_within) %>% 
  st_drop_geometry() %>%
  group_by(uniqueID) %>%
  summarize(Crimecount = n()) %>%
    left_join(heroin20_net, . ) %>%
    st_sf() %>%
  mutate(Crimecount = ifelse(is.na(Crimecount), 0, Crimecount))


heroin20_net <- heroin20_net %>% 
  st_join(complaint20, ., join=st_within) %>% 
  st_drop_geometry() %>%
  group_by(uniqueID) %>%
  summarize(Complaintscount = n()) %>%
    left_join(heroin20_net, . ) %>%
    st_sf() %>%
  mutate(Complaintscount = ifelse(is.na(Complaintscount), 0, Complaintscount))




```



```{r}

# running poission
poissionTrain <- glm(countHeroin ~ Crimecount + Complaintscount, family = "poisson",
                      data = heroinTrain_net)

Prediction19 <-
  mutate(heroin19_net, Prediction = predict(poissionTrain, heroin19_net, type = "response")) %>% 
  mutate(Error = countHeroin - Prediction) %>% 
  mutate(MAE = mean(abs(Error)))


Prediction20 <-
  mutate(heroin20_net, Prediction = predict(poissionTrain, heroin20_net, type = "response")) %>% 
  mutate(Error = countHeroin - Prediction) %>% 
  mutate(MAE = mean(abs(Error)))
```





```{r}

library(viridis)

options(scipen=999)
ggplot() + 
  geom_sf(data = cincinnati, fill = "black") +
  stat_density2d(data = data.frame(st_coordinates(heroin)), 
                 aes(X, Y, fill = after_stat(level), alpha = after_stat(level)),
                 size = 0.01, bins = 60, geom = 'polygon') +
  scale_fill_viridis(option = "magma", name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = "none") +
  labs(title = "Density of Heroin Overdose") +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks =element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.subtitle = element_text(size = 9,face = "italic"),
        plot.title = element_text(size = 12, face = "bold"),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth=0.8)
        )

```


```{r}

heroin_net <- 
  dplyr::select(heroin) %>%  
  mutate(countHeroin = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countHeroin = replace_na(countHeroin, 0),
         uniqueID = 1:n(), 
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

ggplot() +
  geom_sf(data = heroin_net, aes(fill = countHeroin), color = NA) +
  scale_fill_viridis(option = "magma", name = "Robbery Counts") +
  labs(title = "Count of Robberies for the Fishnet") +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks =element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.subtitle = element_text(size = 9,face = "italic"),
        plot.title = element_text(size = 12, face = "bold"),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth=0.8)
        )
```

```{r local moran I}
library(spdep)
library(sfdep)

neigh_nbs <- heroin_net %>% 
  mutate(
    nb = st_contiguity(geometry),  # neighbors share border    
    wt = st_weights(nb), # row-standardized weights              
    neigh_lag = st_lag(countHeroin, nb, wt)  # calculate spatial lag of mean sale price
  )

gi_hot_spots <- neigh_nbs %>% 
  mutate(Gi = local_g_perm(countHeroin, nb, wt, nsim = 999)) %>% 
  unnest(Gi) 



gi_hot_spots <- gi_hot_spots %>%  
  dplyr::select(gi, p_folded_sim, uniqueID) |> 
  mutate(
    classification = case_when(
      # Classify based on the following criteria:
      gi > 0 & p_folded_sim <= 0.01 ~ "Very hot",
      gi > 0 & p_folded_sim <= 0.05 ~ "Hot",
      gi > 0 & p_folded_sim <= 0.1 ~ "Somewhat hot",
      gi < 0 & p_folded_sim <= 0.01 ~ "Very cold",
      gi < 0 & p_folded_sim <= 0.05 ~ "Cold",
      gi < 0 & p_folded_sim <= 0.1 ~ "Somewhat cold",
      TRUE ~ "Insignificant"
    ),    # Convert 'classification' into a factor for easier plotting
    classification = factor(
      classification,
      levels = c("Very hot", "Hot", "Somewhat hot",
                 "Insignificant",
                 "Somewhat cold", "Cold", "Very cold")
    )
  )

gi_hot_spots %>% 
  ggplot() + 
  geom_sf(aes(fill = classification), color = "black", lwd = 0.1) +
    scale_fill_brewer(type = "div", palette = 5) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

```{r}

heroin_net <- gi_hot_spots %>% 
  mutate(hotspot = ifelse(classification %in% c("Very hot", "Hot", "Somewhat hot"), 1, 0)) %>% 
  st_drop_geometry() %>% 
  left_join(heroin_net, ., by = "uniqueID")

```


```{r}


# adding crime to net
heroin_net <- heroin_net %>% 
  st_join(crime, ., join=st_within) %>% 
  st_drop_geometry() %>%
  group_by(uniqueID) %>%
  summarize(Crimecount = n()) %>%
    left_join(heroin_net, . ) %>%
    st_sf() %>%
  mutate(Crimecount = ifelse(is.na(Crimecount), 0, Crimecount))

# add 311 to net
heroin_net <- heroin_net %>% 
  st_join(complaints, ., join=st_within) %>% 
  st_drop_geometry() %>%
  group_by(uniqueID) %>%
  summarize(Complaintscount = n()) %>%
    left_join(heroin_net, . ) %>%
    st_sf() %>%
  mutate(Complaintscount = ifelse(is.na(Complaintscount), 0, Complaintscount))

net_centroid <- st_centroid(heroin_net)
library(FNN)
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

# add hospitals
heroin_net <- heroin_net %>% 
  left_join(net_centroid %>% 
              mutate(hospital.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(hospitals), 1)*0.3048) %>% 
              st_drop_geometry() %>% dplyr::select(hospital.nn, uniqueID), 
            by = "uniqueID")

# add rehab center
heroin_net <- heroin_net %>% 
  left_join(net_centroid %>% 
              mutate(rehab.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(rehab), 1)*0.3048) %>% 
              st_drop_geometry() %>% dplyr::select(rehab.nn, uniqueID), 
            by = "uniqueID")

# add pharmacy
heroin_net <- heroin_net %>% 
  left_join(net_centroid %>% 
              mutate(pharm.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(pharmacy), 1)*0.3048) %>% 
              st_drop_geometry() %>% dplyr::select(pharm.nn, uniqueID), 
            by = "uniqueID")


# add gas station
heroin_net <- heroin_net %>% 
  left_join(net_centroid %>% 
              mutate(fuel.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(fuel), 2)*0.3048) %>% 
              st_drop_geometry() %>% dplyr::select(fuel.nn, uniqueID), 
            by = "uniqueID")

# add fast food restaurant
heroin_net <- heroin_net %>% 
  left_join(net_centroid %>% 
              mutate(fast.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(fastfood), 2)*0.3048) %>% 
              st_drop_geometry() %>% dplyr::select(fast.nn, uniqueID), 
            by = "uniqueID")

# add parks
heroin_net <- heroin_net %>% 
  left_join(net_centroid %>% 
              mutate(parks.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(parks), 2)*0.3048) %>% 
              st_drop_geometry() %>% dplyr::select(parks.nn, uniqueID), 
            by = "uniqueID")


# add demographic vars
heroin_net <- heroin_net %>% 
  left_join(net_centroid %>% 
              select(uniqueID) %>% 
              st_intersection(cincinnati20) %>% 
              st_drop_geometry() %>% select(-GEOID), by = "uniqueID") %>% 
  filter(is.na(pop25_54) == FALSE)
```


```{r logistic without cv}

kitchensink <- glm(hotspot ~ .,
                  data=heroin_net %>% 
                    st_drop_geometry() %>% 
                    dplyr::select(hotspot, Crimecount, Complaintscount, pharm.nn, hospital.nn, fast.nn, fuel.nn, rehab.nn, parks.nn, pop25_54, MedHHInc, MF_ratio, race_ratio, pctPoverty, pctBachelor), family="binomial" (link="logit"))

summary(kitchensink)
```


```{r logistic with CV}

logitCV <- function(dataset, id, dependentVariable, indVariables) {
  
  allPredictions <- data.frame()
  cvID_list <- unique(dataset[[id]])

  for (i in cvID_list) {
    
    thisFold <- i
    cat("This hold out fold is", thisFold, "\n")
    
    fold.train <- filter(dataset, dataset[[id]] != thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, all_of(indVariables),
                    all_of(dependentVariable))
    fold.test  <- filter(dataset, dataset[[id]] == thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, all_of(indVariables),
                    all_of(dependentVariable))
    
    form_parts <- paste0(dependentVariable, " ~ ", paste0(indVariables, collapse = "+"))
    form <- as.formula(form_parts)
    regression <- glm(form, data = fold.train %>%
                        dplyr::select(-geometry, -id), family="binomial" (link="logit"))
    
    thisPrediction <-
      mutate(fold.test, Prediction = predict(regression, fold.test, type = "response"))
    
    allPredictions <-
      rbind(allPredictions, thisPrediction)
    
  }
  return(st_sf(allPredictions))
}

```


```{r}

reg.vars <- c("Crimecount", "Complaintscount", "pharm.nn", "hospital.nn", "fast.nn", "fuel.nn", "rehab.nn", "parks.nn", "pop25_54", "MedHHInc", "MF_ratio", "race_ratio", "pctPoverty", "pctBachelor")


kitchensink.logit.cv <- logitCV (
  dataset = heroin_net,
  id = "cvID",
  dependentVariable = "hotspot",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = cvID, hotspot, Prediction, geometry)

kitchensink.logit.cv <- kitchensink.logit.cv %>% 
  mutate(PredictionCat  = as.factor(ifelse(kitchensink.logit.cv$Prediction > 0.5 , 1, 0))) %>% 
  mutate(hotspot = as.factor(hotspot))

caret::confusionMatrix(kitchensink.logit.cv$hotspot, kitchensink.logit.cv$PredictionCat, 
                       positive = "1")

```


```{r PCA}

library(stats)

pca_heroin <- heroin_net %>% 
  st_drop_geometry() %>% 
  select(Crimecount, Complaintscount, pharm.nn, hospital.nn, fast.nn, fuel.nn, rehab.nn, parks.nn, pop25_54, MedHHInc, MF_ratio, race_ratio, pctPoverty, pctBachelor)

pca_heroin <- na.omit(pca_heroin)
pc <- prcomp(pca_heroin,
             center = TRUE,
            scale. = TRUE)
summary(pc)
print(pc)

library(factoextra)
fviz_eig(pc, addlabels = TRUE)
fviz_pca_var(pc, col.var = "black")

fviz_cos2(pc, choice = "var", axes = 1:7)


 

```

```{r}

pca_net <- heroin_net %>% 
  mutate(pc1 = Crimecount * pc$rotation[1] + Complaintscount * pc$rotation[2] + pharm.nn * pc$rotation[3] + hospital.nn * pc$rotation[4] + fast.nn * pc$rotation[5] +  fuel.nn * pc$rotation[6] + rehab.nn * pc$rotation[7] + parks.nn *  pc$rotation[8] + pop25_54 * pc$rotation[9] + MedHHInc * pc$rotation[10] + MF_ratio * pc$rotation[11] + race_ratio * pc$rotation[12] +  pctPoverty * pc$rotation[13] + pctBachelor * pc$rotation[14]) %>% 
  mutate(pc2 = Crimecount * pc$rotation[15] + Complaintscount * pc$rotation[16] + pharm.nn * pc$rotation[17] + hospital.nn * pc$rotation[18] + fast.nn * pc$rotation[19] +  fuel.nn * pc$rotation[20] + rehab.nn * pc$rotation[21] + parks.nn *  pc$rotation[22] + pop25_54 * pc$rotation[23] + MedHHInc * pc$rotation[24] + MF_ratio * pc$rotation[25] + race_ratio * pc$rotation[26] +  pctPoverty * pc$rotation[27] + pctBachelor * pc$rotation[28]) %>%
  mutate(pc3 = Crimecount * pc$rotation[29] + Complaintscount * pc$rotation[30] + pharm.nn * pc$rotation[31] + hospital.nn * pc$rotation[32] + fast.nn * pc$rotation[33] +  fuel.nn * pc$rotation[34] + rehab.nn * pc$rotation[35] + parks.nn *  pc$rotation[36] + pop25_54 * pc$rotation[37] + MedHHInc * pc$rotation[38] + MF_ratio * pc$rotation[39] + race_ratio * pc$rotation[40] +  pctPoverty * pc$rotation[41] + pctBachelor * pc$rotation[42]) %>%
  mutate(pc4 = Crimecount * pc$rotation[43] + Complaintscount * pc$rotation[44] + pharm.nn * pc$rotation[45] + hospital.nn * pc$rotation[46] + fast.nn * pc$rotation[47] +  fuel.nn * pc$rotation[48] + rehab.nn * pc$rotation[49] + parks.nn *  pc$rotation[50] + pop25_54 * pc$rotation[51] + MedHHInc * pc$rotation[52] + MF_ratio * pc$rotation[53] + race_ratio * pc$rotation[54] +  pctPoverty * pc$rotation[55] + pctBachelor * pc$rotation[56]) %>%
  mutate(pc5 = Crimecount * pc$rotation[57] + Complaintscount * pc$rotation[58] + pharm.nn * pc$rotation[59] + hospital.nn * pc$rotation[60] + fast.nn * pc$rotation[61] +  fuel.nn * pc$rotation[62] + rehab.nn * pc$rotation[63] + parks.nn *  pc$rotation[64] + pop25_54 * pc$rotation[65] + MedHHInc * pc$rotation[66] + MF_ratio * pc$rotation[67] + race_ratio * pc$rotation[68] +  pctPoverty * pc$rotation[69] + pctBachelor * pc$rotation[70]) %>%
  mutate(pc6 = Crimecount * pc$rotation[71] + Complaintscount * pc$rotation[72] + pharm.nn * pc$rotation[73] + hospital.nn * pc$rotation[74] + fast.nn * pc$rotation[75] +  fuel.nn * pc$rotation[76] + rehab.nn * pc$rotation[77] + parks.nn *  pc$rotation[78] + pop25_54 * pc$rotation[79] + MedHHInc * pc$rotation[80] + MF_ratio * pc$rotation[81] + race_ratio * pc$rotation[82] +  pctPoverty * pc$rotation[83] + pctBachelor * pc$rotation[84]) %>%
  mutate(pc7 = Crimecount * pc$rotation[85] + Complaintscount * pc$rotation[86] + pharm.nn * pc$rotation[87] + hospital.nn * pc$rotation[88] + fast.nn * pc$rotation[89] +  fuel.nn * pc$rotation[90] + rehab.nn * pc$rotation[91] + parks.nn *  pc$rotation[92] + pop25_54 * pc$rotation[93] + MedHHInc * pc$rotation[94] + MF_ratio * pc$rotation[95] + race_ratio * pc$rotation[96] +  pctPoverty * pc$rotation[97] + pctBachelor * pc$rotation[98])
  

```


```{r}

reg.vars <- c("pc1", "pc2", "pc3", "pc4", "pc5", "pc6", "pc7")


pca.logit.cv <- logitCV (
  dataset = pca_net,
  id = "cvID",
  dependentVariable = "hotspot",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = cvID, hotspot, Prediction, geometry)

pca.logit.cv <- pca.logit.cv %>% 
  mutate(PredictionCat  = as.factor(ifelse(pca.logit.cv$Prediction > 0.5 , 1, 0))) %>% 
  mutate(hotspot = as.factor(hotspot))

caret::confusionMatrix(pca.logit.cv$hotspot, pca.logit.cv$PredictionCat, 
                       positive = "1")
```



```{r}
library(plotROC)
ggplot(pca.logit.cv, aes(d = as.numeric(hotspot), m = Prediction)) +
  geom_roc(n.cuts = 50, labels = FALSE, colour = "#2E4756") +
    labs(title = "ROC Curve for Improved Model") +
  style_roc(theme = theme_grey) +
  geom_abline(slope = 1, intercept = 0, size = 1.5, color = "#DBC2CF") +
  theme(plot.subtitle = element_text(size = 9,face = "italic"),
        plot.title = element_text(size = 12, face = "bold"), 
        axis.text.x=element_text(size=8),
        axis.text.y=element_text(size=8), 
        axis.title=element_text(size=9), 
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth =0.8))
```

```{r}
library(classInt)
library(raster)
library(spatstat.explore)
rob_ppp <- as.ppp(st_coordinates(heroin), W = st_bbox(heroin_net))
rob_KD.1000 <- spatstat.explore::density.ppp(rob_ppp, 1000)
rob_KD.df <- data.frame(rasterToPoints(mask(raster(rob_KD.1000), as(cincinnati20, 'Spatial'))))


ggplot(data=rob_KD.df, aes(x=x, y=y)) +
  geom_raster(aes(fill=layer)) + 
  coord_sf(crs=st_crs(heroin_net)) + 
  scale_fill_viridis(option = "magma", name="Density") +
  labs(title = "Kernel Density of Heroin 1000ft Radii") +
    theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks =element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.subtitle = element_text(size = 9,face = "italic"),
        plot.title = element_text(size = 12, face = "bold"),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, size=0.8)
        )
```

```{r}

rob_KDE_sum <- as.data.frame(rob_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(heroin_net)) %>%
  aggregate(., heroin_net, mean) 

kde_breaks <- classIntervals(rob_KDE_sum$value, 
                             n = 2, "fisher")

rob_KDE_sf <- rob_KDE_sum %>%
  mutate(label = "Kernel Density",
         Risk_Category = classInt::findCols(kde_breaks),
         Risk_Category = case_when(
           Risk_Category == 2 ~ "High",
           Risk_Category == 1 ~ "Low"))

rob_risk_sf <-
  pca.logit.cv %>%
  mutate(label = "Logistic Regression",
         Risk_Category = case_when(
           PredictionCat == 1 ~ "High",
           PredictionCat == 0 ~ "Low")) %>% 
  dplyr::select(PredictionCat, label, Risk_Category) %>% 
  rename(value = PredictionCat)


rbind(rob_KDE_sf, rob_risk_sf) %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category, -geometry) %>%
  ggplot() +
    geom_sf(aes(fill = Risk_Category), colour = NA) +
    facet_wrap(~label, ) +
    scale_fill_viridis(option = "magma", discrete = TRUE, name = "Risk Category") +
    #labs(title="Comparison of Kernel Density and Risk Predictions",
         #subtitle="2021 robbery risk predictions; 2022 robberies") + 
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks =element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.subtitle = element_text(size = 9,face = "italic"),
        plot.title = element_text(size = 12, face = "bold"),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, size=0.8)
        )
```



```{r}

scale_values <- function(x){(x-min(x))/(max(x)-min(x))}

score <- heroin_net %>% 
  mutate(MedHHInc = ifelse(is.na(MedHHInc), 0, MedHHInc)) %>% 
  mutate(scl_crime = scale_values(Crimecount),
         scl_complaints = scale_values(Complaintscount),
         scl_pharm = scale_values(pharm.nn),
         scl_hospital = scale_values(hospital.nn),
         scl_rehab = scale_values(rehab.nn),
         scl_park = scale_values(parks.nn),
         scl_fuel = scale_values(fuel.nn),
         scl_fast = scale_values(fast.nn),
         scl_pop = scale_values(pop25_54),
         scl_income = scale_values(MedHHInc),
         scl_gender = scale_values(MF_ratio),
         scl_race = scale_values(race_ratio),
         scl_poverty = scale_values(pctPoverty),
         scl_edu = scale_values(pctBachelor)) %>% 
  mutate(scl_race_re = 0 - scl_race + 1,
         scl_edu_re = 0 - scl_edu + 1,
         scl_income_re = 0 - scl_income + 1, 
         scl_fuel_re = 0 - scl_fuel + 1,
         scl_fast_re = 0 - scl_fast + 1,
         scl_park_re = 0 - scl_park + 1,
         scl_pharm_re = 0 - scl_pharm + 1) %>% 
  mutate(score = 0.1 * (scl_hospital + scl_crime + scl_complaints +  scl_gender + scl_income_re + scl_fuel_re + scl_fast_re +  scl_park_re) + 0.05 * (scl_race_re +  scl_pharm_re) + 0.025*(scl_pop +scl_edu_re + scl_poverty +  scl_rehab))
  

```


Predictor Variables so far: 
- hospitals 0.1
- crime rate (2016-2018) 0.1
- 311 complaints (2016-2018) 0.1
- population between 25-54 0.025
- gender ratio 0.1
- *white over minority ratio 0.05*
- *education 0.025*
- *median household income 0.1*
- poverty 0.025
- *gas station 0.1*
- *fast food restaurant 0.1*
- *public parks 0.1*
- *pharmacies 0.05*
- rehabilitation centers 0.025


```{r}
library(mapview)

mapview(score, zcol = "score", fill = "score", alpha = 0.3) 
```





```{r poisson regresssion}

reg.vars <- c("Crimecount", "Complaintscount", "pharm.nn", "hospital.nn", "fast.nn", "fuel.nn", "rehab.nn", "parks.nn", "pop25_54", "MedHHInc", "MF_ratio", "race_ratio", "pctPoverty", "pctBachelor")

kitchensink.cv <- crossValidate(
  dataset = heroin_net,
  id = "cvID",
  dependentVariable = "countHeroin",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = cvID, countHeroin, Prediction, geometry)

kitchensink.summary <- kitchensink.cv %>% 
  mutate(Error = Prediction - countHeroin) %>% 
  mutate(Mean_Error = mean(Error, na.rm=T))
  
```


